---
title: "1 A Layered Grammar of Graphics: ggplot2 methods"
author: "Xiao Fudi"
date: "April 13, 2024"
date-modified: "last-modified"
execute:
  Eval: true
  Echo: true
  Warning: false
  Freeze: true
---


# **Overview**

In this chapter, we will learn the basic principles and essential components of ggplot2, and gain hands-on experience on using these components to plot statistical graphics based on the principle of Layered Grammar of Graphics

By then end of this chapter, we will be able to apply the essential graphical elements provided by ggplot2 to create elegant and yet functional statistical graphics.

# **1 Getting Started**

### **1.1 Loading R package**

```{r}
pacman::p_load(tidyverse)
```

### **1.2 Importing Data**

```{r}
exam_data <- read_csv("data/Exam_data.csv")

```

## **1.3 Introducing ggplot**

ggplot2 is an R package for creating data-driven graphics based on The Grammar of Graphics. It is also part of the tidyverse family specially designed for visual exploration and communication.

For more information, please visit [ggplot2 link](https://ggplot2.tidyverse.org/)

## **1.3.1 R Graphics VS ggplot**

First, let us compare how R Graphics, the core graphical functions of Base R and ggplot plot a simple histogram.

```{r}
hist(exam_data$MATHS, col = '#5e82c9')
```

```{r}
ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=10, 
                 boundary = 100,
                 color="black", 
                 fill='#5e82c9') +
  ggtitle("Distribution of Maths scores")
```

## **1.4 Grammar of Graphics**

Grammar of Graphics is a general scheme for data visualization which breaks up graphs into semantic components such as scales and layers. It was introduced by Leland Wilkinson (1999) **Grammar of Graphics**, Springer. The grammar of graphics is an answer to a question:

What is a statistical graphic?

In the nutshell, Grammar of Graphics defines the rules of structuring mathematical and aesthetic elements into a meaningful graph.

There are two principles in Grammar of Graphics, they are:

-   Graphics = distinct layers of grammatical elements

-   Meaningful plots through aesthetic mapping

![Reference: Hadley Wickham (2010) [“A layered grammar of graphics.”](https://vita.had.co.nz/papers/layered-grammar.html) *Journal of Computational and Graphical Statistics*, vol. 19, no. 1, pp. 3–28.](images/clipboard-1417645947.png)

A short description of each building block are as follows:

-   **Data**: The dataset being plotted.

-   **Aesthetics** take attributes of the data and use them to influence visual characteristics, such as position, colours, size, shape, or transparency.

-   **Geometrics**: The visual elements used for our data, such as point, bar or line.

-   **Facets** split the data into subsets to create multiple variations of the same graph (paneling, multiple plots).

-   **Statistics**, statiscal transformations that summarise data (e.g. mean, confidence intervals).

-   **Coordinate systems** define the plane on which data are mapped on the graphic.

-   **Themes** modify all non-data components of a plot, such as main title, sub-title, y-aixs title, or legend background.

## **1.5 Essential Grammatical Elements in ggplot2: data**

```         
Let us call the `ggplot()` function using the code chunk on the right.
```

```{r}
ggplot(data=exam_data)
```

-   A blank canvas appears.

-   `ggplot()` initializes a ggplot object.

-   The *data* argument defines the dataset to be used for plotting.

-   If the dataset is not already a data.frame, it will be converted to one by `fortify()`.

## 1.6 Essential Grammatical Elements in ggplot2: Aesthetic mappings

