---
title: "Hands-on_Ex03(a)"
---

# **3. Programming Interactive Data Visualisation with R**

## **3.1 Learning Outcome**

In this hands-on exercise, we will learn how to create interactive data visualisation by using functions provided by **ggiraph** and **plotlyr** packages.

## **3.2 Getting Started**

The code chunk below will be used to accomplish the task.

-   [**ggiraph**](https://davidgohel.github.io/ggiraph/) for making ‘ggplot’ graphics interactive.

-   [**plotly**](https://plotly.com/r/), R library for plotting interactive statistical graphs.

-   [**DT**](https://rstudio.github.io/DT/) provides an R interface to the JavaScript library [DataTables](https://datatables.net/) that create interactive table on html page.

-   [**tidyverse**](https://www.tidyverse.org/), a family of modern R packages specially designed to support data science, analysis and communication task including creating static statistical graphs.

-   [**patchwork**](https://patchwork.data-imaginist.com/) for combining multiple ggplot2 graphs into one figure.

    ```{r}
    pacman::p_load(ggiraph, plotly, 
                   patchwork, DT, tidyverse) 
    ```

## **3.3 Importing Data**

*Exam_data.csv* provided will be used. Using [*read_csv()*](https://readr.tidyverse.org/reference/read_delim.html) of **readr** package, import *Exam_data.csv* into R.

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

## **3.4 Interactive Data Visualisation - ggiraph methods**

[ggiraph](https://davidgohel.github.io/ggiraph/) is an htmlwidget and a ggplot2 extension. It allows ggplot graphics to be interactive.

Interactive is made with [**ggplot geometries**](https://davidgohel.github.io/ggiraph/reference/#section-interactive-geometries) that can understand three arguments:

-   **Tooltip**: a column of data-sets that contain tooltips to be displayed when the mouse is over elements.

-   **Onclick**: a column of data-sets that contain a JavaScript function to be executed when elements are clicked.

-   **Data_id**: a column of data-sets that contain an id to be associated with elements.

### **3.4.1 Tooltip effect with *tooltip* aesthetic**

Below shows a typical code chunk to plot an interactive statistical graph by using **ggiraph** package. [`girafe()`](https://davidgohel.github.io/ggiraph/reference/girafe.html) of **ggiraph** will be used to create an interactive svg object.

```{r}
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = ID),
    stackgroups = TRUE, 
    binwidth = 1, 
    method = "histodot") +
  scale_y_continuous(NULL, 
                     breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
)
```

## **3.5 Interactivity**

By hovering the mouse pointer on an data point of interest, the student’s ID will be displayed.

```{r}
exam_data$tooltip <- c(paste0(     
  "Name = ", exam_data$ID,         
  "\n Class = ", exam_data$CLASS)) 

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = exam_data$tooltip), 
    stackgroups = TRUE,
    binwidth = 1,
    method = "histodot") +
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 8,
  height_svg = 8*0.618
)
```

## **3.6 Interactivity**

### **3.6.1 Customising Tooltip style**

Code chunk below uses [`opts_tooltip()`](https://davidgohel.github.io/ggiraph/reference/opts_tooltip.html) of **ggiraph** to customize tooltip rendering by add css declarations.

```{r}
tooltip_css <- "background-color:white; #<<
font-style:bold; color:black;" #<<

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(tooltip = ID),                   
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(    #<<
    opts_tooltip(    #<<
      css = tooltip_css)) #<<
)                                        
```

### **3.6.2 Displaying statistics on tooltip**

Code chunk below shows an advanced way to customise tooltip. In this example, a function is used to compute 90% confident interval of the mean. The derived statistics are then displayed in the tooltip.

```{r}
tooltip <- function(y, ymax, accuracy = .01) {
  mean <- scales::number(y, accuracy = accuracy)
  sem <- scales::number(ymax - y, accuracy = accuracy)
  paste("Mean maths scores:", mean, "+/-", sem)
}

gg_point <- ggplot(data=exam_data, 
                   aes(x = RACE),
) +
  stat_summary(aes(y = MATHS, 
                   tooltip = after_stat(  
                     tooltip(y, ymax))),  
    fun.data = "mean_se", 
    geom = GeomInteractiveCol,  
    fill = "light blue"
  ) +
  stat_summary(aes(y = MATHS),
    fun.data = mean_se,
    geom = "errorbar", width = 0.2, size = 0.2
  )

girafe(ggobj = gg_point,
       width_svg = 8,
       height_svg = 8*0.618)
```

### **3.6.3 Hover effect with *data_id* aesthetic**

Code chunk below shows the second interactive feature of ggiraph, namely `data_id`.

```{r}
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(           
    aes(data_id = CLASS),             
    stackgroups = TRUE,               
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618                      
)                                        
```

Note that the default value of the hover css is *hover_css = “fill:orange;”*.

### **3.6.4 Styling hover effect**

In the code chunk below, css codes are used to change the highlighting effect.

```{r}
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(data_id = CLASS),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(                        
    opts_hover(css = "fill: #202020;"),  
    opts_hover_inv(css = "opacity:0.2;") 
  )                                        
)                                        
```

::: callout-tip
Note: Different from previous example, in this example the ccs customisation request are encoded directly.
:::

### **3.6.5 Combining tooltip and hover effect**

There are time that we want to combine tooltip and hover effect on the interactive statistical graph as shown in the code chunk below.

```{r}
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(tooltip = CLASS, 
        data_id = CLASS),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(                        
    opts_hover(css = "fill: #202020;"),  
    opts_hover_inv(css = "opacity:0.2;") 
  )                                        
)                                        
```

Interactivity: Elements associated with a *data_id* (i.e CLASS) will be highlighted upon mouse over. At the same time, the tooltip will show the CLASS.

### **3.6.6 Click effect with onclick**

`onclick` argument of ggiraph provides hotlink interactivity on the web.

The code chunk below shown an example of `onclick`.

```{r}
exam_data$onclick <- sprintf("window.open(\"%s%s\")",
"https://www.moe.gov.sg/schoolfinder?journey=Primary%20school",
as.character(exam_data$ID))

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(onclick = onclick),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618)                                        
```

::: callout-warning
Note that click actions must be a string column in the dataset containing valid javascript instructions.
:::

### **3.6.7 Coordinated Multiple Views with ggiraph**

Coordinated multiple views methods has been implemented in the data visualisation below.

In order to build a coordinated multiple views as shown in the example above, the following programming strategy will be used:

1.  Appropriate interactive functions of **ggiraph** will be used to create the multiple views.

2.  *patchwork* function of [patchwork](https://patchwork.data-imaginist.com/) package will be used inside girafe function to create the interactive coordinated multiple views.

    ```{r}
    p1 <- ggplot(data=exam_data, 
           aes(x = MATHS)) +
      geom_dotplot_interactive(              
        aes(data_id = ID),              
        stackgroups = TRUE,                  
        binwidth = 1,                        
        method = "histodot") +  
      coord_cartesian(xlim=c(0,100)) + 
      scale_y_continuous(NULL,               
                         breaks = NULL)

    p2 <- ggplot(data=exam_data, 
           aes(x = ENGLISH)) +
      geom_dotplot_interactive(              
        aes(data_id = ID),              
        stackgroups = TRUE,                  
        binwidth = 1,                        
        method = "histodot") + 
      coord_cartesian(xlim=c(0,100)) + 
      scale_y_continuous(NULL,               
                         breaks = NULL)

    girafe(code = print(p1 + p2), 
           width_svg = 6,
           height_svg = 3,
           options = list(
             opts_hover(css = "fill: #202020;"),
             opts_hover_inv(css = "opacity:0.2;")
             )
           ) 
    ```

## **3.7 Interactive Data Visualisation - plotly methods!**

Plotly’s R graphing library create interactive web graphics from **ggplot2** graphs and/or a custom interface to the (MIT-licensed) JavaScript library [**plotly.js**](https://plotly.com/javascript/) inspired by the grammar of graphics.

There are two ways to create interactive graph by using plotly, they are:

-   by using *plot_ly()*, and

-   by using *ggplotly()*

### **3.7.1 Creating an interactive scatter plot: plot_ly() method**

The tabset below shows an example a basic interactive plot created by using *plot_ly()*.

```{r}
plot_ly(data = exam_data, 
             x = ~MATHS, 
             y = ~ENGLISH)
```

### **3.7.2 Working with visual variable: plot_ly() method**

In the code chunk below, *color* argument is mapped to a qualitative visual variable (i.e. RACE).

```{r}
plot_ly(data = exam_data, 
        x = ~ENGLISH, 
        y = ~MATHS, 
        color = ~RACE)
```

### **3.7.3 Creating an interactive scatter plot: ggplotly() method**

The code chunk below plots an interactive scatter plot by using *ggplotly()*.

```{r}
p <- ggplot(data=exam_data, 
            aes(x = MATHS,
                y = ENGLISH)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
ggplotly(p)
```

### **3.7.4 Coordinated Multiple Views with plotly**

The creation of a coordinated linked plot by using plotly involves three steps:

-   [`highlight_key()`](https://www.rdocumentation.org/packages/plotly/versions/4.9.2/topics/highlight_key) of **plotly** package is used as shared data.

-   two scatterplots will be created by using ggplot2 functions.

-   lastly, [*subplot()*](https://plotly.com/r/subplots/) of **plotly** package is used to place them next to each other side-by-side.

```{r}
d <- highlight_key(exam_data)
p1 <- ggplot(data=d, 
            aes(x = MATHS,
                y = ENGLISH)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))

p2 <- ggplot(data=d, 
            aes(x = MATHS,
                y = SCIENCE)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
subplot(ggplotly(p1),
        ggplotly(p2))
```

Thing to learn from the code chunk:

-   `highlight_key()` simply creates an object of class [crosstalk::SharedData](https://rdrr.io/cran/crosstalk/man/SharedData.html).

-   Visit this [link](https://rstudio.github.io/crosstalk/) to learn more about crosstalk

## **3.8 Interactive Data Visualisation - crosstalk methods!**

### **3.8.1 Interactive Data Table: DT package**

-   A wrapper of the JavaScript Library [DataTables](https://datatables.net/)

-   Data objects in R can be rendered as HTML tables using the JavaScript library ‘DataTables’ (typically via R Markdown or Shiny).

    ```{r}
    DT::datatable(exam_data, class= "compact")
    ```

### **3.8.2 Linked brushing: crosstalk method**

```{r}
d <- highlight_key(exam_data) 
p <- ggplot(d, 
            aes(ENGLISH, 
                MATHS)) + 
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))

gg <- highlight(ggplotly(p),        
                "plotly_selected")  

crosstalk::bscols(gg,               
                  DT::datatable(d), 
                  widths = 5)        
```

Things to learn from the code chunk:

-   *highlight()* is a function of **plotly** package. It sets a variety of options for brushing (i.e., highlighting) multiple plots. These options are primarily designed for linking multiple plotly graphs, and may not behave as expected when linking plotly to another htmlwidget package via crosstalk. In some cases, other htmlwidgets will respect these options, such as persistent selection in leaflet.

-   *bscols()* is a helper function of **crosstalk** package. It makes it easy to put HTML elements side by side. It can be called directly from the console but is especially designed to work in an R Markdown document. **Warning:** This will bring in all of Bootstrap!.
